---
title: "SET Rate Calculations"
date: "`r Sys.Date()`"
output: word_document
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = TRUE)
```

In this report, the user may have chosen to exclude data associated with certain QA/QC codes. Any values that have been removed are listed in the appropriate section below.    

Also, be aware that linear models are not appropriate for calculating rates of change at all sites. Use discretion when interpreting these results.  


# Setup  



```{r load-libraries}
library(knitr)
library(forcats)
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(nlme)
library(broom)
library(here)
library(flextable)
library(readxl)
library(lubridate)
library(leaflet)
# load functions
source(here::here('R_scripts', 'sourced', '000_functions.R'))
# load user options
source(here::here('R_scripts', 'sourced', '001_user_options.R'))
```


## Read in data and metadata  

This reads in the long dataset, converted from other formats by earlier scripts. It also converts pin heights to mm if they weren't already in those units.  

If the user specified custom start and/or end dates in the file `metadata\user_defined_inputs.xlsx`, the data frame is trimmed in this step.  

```{r read-data, warning = TRUE}
# find the folder with processed data
path <- here::here('data', 'processed')

# in that folder, find the name of the file(s) that ends with 'set_processed.csv'
filelist <- dir(path, pattern = "set_processed.csv$")


# generate warnings if the folder is empty; if there are multiple matching files, select the first one
if (length(filelist) == 0) stop("There are no files of the correct name/format (---set_QC.csv) in your processed data folder.")

if (length(filelist) > 1) {
    warning("There is more than one file of the correct format (---set_QC.csv) in your data folder. The first file alphabetically will be used.")
    filelist <- filelist[1]
}



# generate the full path to the file; read it in and get pin heights to mm
filename <- paste0(path, "/", filelist)
dat <- read_csv(filename)
dat <- height_to_mm(dat)



# get rid of any placeholders; make sure set_id is a factor;
# paste year, month, and day together into "date"
dat <- dat %>%
    mutate(date = lubridate::ymd(paste(year, month, day))) %>% 
    filter(!is.na(date),
           !is.na(pin_number)) %>%
    mutate(set_id = as.factor(set_id),
           arm_position = as.character(arm_position))


# trim the dataset based on user-specified dates
if(!is.na(opts_long$custom_start)){
  start <- lubridate::ymd(opts_long$custom_start)
  dat <- dat %>% 
    filter(date >= start)
}
if(!is.na(opts_long$custom_end)){
  end <- lubridate::ymd(opts_long$custom_end)
  dat <- dat %>% 
    filter(date <= end)
}
```


## QA/QC codes  

Use the *user_defined_inputs* spreadsheet to find qa/qc codes that the user wants to have removed (designated in that sheet with "-3" to be consistent with the SWMP "reject" flagging). Report those values and then turn them into NAs.

Note that this code will look for an EXACT match. e.g., if "LHE" is specified for removal, only values with "LHE" will be removed. "LHE CB" and "CRM LHE" will *not* be removed. To remove a combination of codes, a new line may need to be created in the *qaqc_codes* tab of the input spreadsheet.  

```{r qc-remove}
if (excl_exist) {
  # find the rows to exclude - either arm or pin codes match
  pin_to_excl <- which(dat$qaqc_code %in% codes_to_exclude)
  arm_to_excl <- which(dat$arm_qaqc_code %in% codes_to_exclude)
  
  # make a data frame of the excluded ones and print it out
  to_excl <- c(pin_to_excl, arm_to_excl)
  excluded <- dat[to_excl, ]
  
  
  excluded %>% 
    select(set_id, date, 
           arm = arm_position, 
           arm_code = arm_qaqc_code, 
           pin_number, qaqc_code, pin_height) %>% 
    knit_print()
  

  # turn the heights into NAs
  dat$pin_height[to_excl] <- NA 
  
  # clean up
  # rm(list = c("excluded", "arm_to_excl", "pin_to_excl", "to_excl"))
}
```


### File Matching checks  

This analysis was run on `r filelist[1]` on `r Sys.Date()`.  


Make sure a metadata file exists:  

```{r read-metadata, warning = TRUE}
# find the folder with metadata
path <- here::here('metadata')

# in that folder, find the name of the file(s) that ends with 'setm.csv'
filelist <- dir(path, pattern = 'set_metadata.xls')

# generate warnings if the folder is empty; if there are multiple matching files, select the first one
if (length(filelist) == 0) stop("There are no files of the correct name/format (---set_metadata.xls or .xlsx) in your metadata folder.")

if (length(filelist) > 1) {
    warning("There is more than one file of the correct format (---set_metadata.xls or .xlsx) in your metadata folder. The first file alphabetically will be used.")
    filelist <- filelist[1]
}

# generate the full path to the file; read it in, clean the names, get rid of empty rows and columns
filename <- paste0(path, "/", filelist)
mdat <- read_excel(filename) %>%
    janitor::clean_names() %>%
    janitor::remove_empty(which = "rows")
```


Make sure the same SET IDs exist in both the data and metadata file:     

```{r verify-setid, warning = TRUE}
# first pull out set_id from both data frames
data_setid <- unique(as.character(dat$set_id))
metadata_setid <- unique(mdat$unique_set_id)

# find set_ids that are in the data, but not in the metadata
dat_not_m <- setdiff(data_setid, metadata_setid)
# find set_ids that are in the metadata, but not in the data
m_not_dat <- setdiff(metadata_setid, data_setid)

if (length(dat_not_m) > 0) {
    toprint <- paste(dat_not_m, collapse = ", ")
    warning(paste0("The following SET IDs exist in your data, but not in your metadata: ", toprint))
}

if (length(m_not_dat) > 0) {
    toprint <- paste(m_not_dat, collapse = ", ")
    warning(paste0("The following SET IDs exist in your metadata, but not in your data: ", toprint))
}

if (length(dat_not_m) + length(m_not_dat) == 0) {
    print("SET IDs match in your data and metadata files.")
}

# go ahead and order the set_ids in dat and unique_set_id in mdat by the metadata's numerical_order, if available
if (!anyNA(mdat$numerical_order)){
  mdat_sub <- select(mdat, unique_set_id, numerical_order) %>% 
    mutate(unique_set_id = factor(unique_set_id))
  dat <- left_join(dat, mdat_sub, by = c("set_id" = "unique_set_id")) %>% 
    mutate(set_id = fct_reorder(set_id, numerical_order)) %>% 
    select(-numerical_order)
  mdat <- mdat %>% 
    mutate(unique_set_id = factor(unique_set_id),
           unique_set_id = fct_reorder(unique_set_id, numerical_order))
}

# cleanup
rm(dat_not_m, m_not_dat, mdat_sub, metadata_setid, data_setid, filelist, filename, path)
```



Read in general sea-level rates sheet and pull out the rate that matches the reserve.  

```{r read-slr}
slr_file <- here::here('metadata', 'slr_rates.csv')
slr_rates <- read_csv(slr_file) %>%
    clean_names() %>%
    janitor::remove_empty(which = c("rows", "cols"))
```

Pull out the information relevant to the current reserve. Assign relevant values to objects and print out the table with information about the site.   

```{r verify-slr, warning = TRUE}
res_to_match <- unique(dat$reserve)

if (res_to_match %in% unique(slr_rates$reserve)) {
    slr_res <- slr_rates %>%
        filter(reserve == res_to_match) %>%
        select(-link)
    slr <- slr_res$slr_rate_mm_yr
    slr_ci <- slr_res$x95_percent_ci
} else {warning("This reserve does not have an entry in the sea level rise rates file. Please check metadata/slr_rates.csv and make sure your reserve is present.")}
```



# Background information  


## Reserve-level  


+  Local rate of sea level change is **`r slr`** +/- **`r slr_ci`** mm/yr.  
+  This rate is reported by `r slr_res$nearest_nwlon_station`, NWLON station number `r as.character(slr_res$nwlon_station_number)` based on data from *`r slr_res$data_start`* to *`r slr_res$data_end`*.
  

## SET-level characteristics  

### setting  


```{r set-info-table}
# print the table
# what to print depends on whether lat/long are to be included

if(opts_long$coords_tech == 1){
  mdat %>% 
    select(unique_set_id, set_type, latitude_dec_deg, 
           longitude_dec_deg, co_dominant_species1) %>% 
    arrange(unique_set_id) %>% 
    setNames(c("SET_ID", "Type", "Lat", "Long", "Main_Veg")) %>% 
    flextable() %>% 
    theme_booktabs() %>% 
    autofit()
} else {
  mdat %>% 
    select(unique_set_id, set_type, co_dominant_species1) %>% 
    arrange(unique_set_id) %>% 
    setNames(c("SET_ID", "Type", "Main_Veg")) %>% 
    flextable() %>% 
    theme_booktabs() %>% 
    autofit()
}
```



### sampling information  

```{r sampling-info-table}
sample_info <- dat %>% 
  group_by(set_id) %>% 
  summarize(first_sampled = min(date),
            last_sampled = max(date),
            years_sampled = as.double(max(date) - min(date)) / 365.25,
            sample_events = length(unique(date))) 

sample_info %>% 
  flextable() %>% 
  theme_booktabs() %>% 
  autofit()
```


Still on wish list to include:  

+  NAVD88 elevation (and year determined)  
+  Distance from closest water body 


***

### Cumulative change snapshot  

Take a look at your overall change since the first reading - make sure the change looks generally linear, and make sure there are no big breaks in the data that could influence the outputs. Output will be generated even if it is not appropriate - **it is up to you to use discretion and make sure a linear model is appropriate for the data!**

```{r, fig.width = opts_long$width_square, fig.height = opts_long$height_square}
dat_cumu <- calc_change_cumu(dat)

# without linear regression
plot_cumu_set(dat_cumu$set, columns = opts_long$multi_cols, smooth = FALSE)

# save it
out_name <- paste("R_output", "figures", "cumu_change_plots", "cumu_change_noLine.png", sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, width = opts_long$width_square, height = opts_long$height_square, units = "in")

```

**The above graph is saved as:**   
_`r out_name`_   



```{r}
all_sets <- unique(dat_cumu$set$set_id)
out_path <- paste("R_output", "figures", "cumu_change_plots", "individual_sets", sep = "/")

for(i in seq_along(all_sets)){
  dat_tmp <- dat_cumu$set[dat_cumu$set$set_id == all_sets[i], ]
  p <- plot_cumu_set(dat_tmp, columns = 1, smooth = FALSE)
  plot_name = paste0("cumu_change_noLine_", all_sets[i], ".png")
  out_name <- paste(out_path, plot_name, sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, p, width = opts_long$width_short, height = opts_long$height_short, units = "in", scale = 0.8)
}
```


Graphs for each SET individually are not shown here but have been saved in _`r out_path`_


```{r, fig.width = opts_long$width_square, fig.height = opts_long$height_square}
# with linear regression
plot_cumu_set(dat_cumu$set, columns = opts_long$multi_cols,
              pointsize = 2.5)

# save it
out_name <- paste("R_output", "figures", "cumu_change_plots", "cumu_change_withLine.png", sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, width = opts_long$width_square, height = opts_long$height_square, units = "in")
```

**The above graph is saved as:**   
_`r out_name`_   


```{r}
out_path <- paste("R_output", "figures", "cumu_change_plots", "individual_sets", sep = "/")

for(i in seq_along(all_sets)){
  dat_tmp <- dat_cumu$set[dat_cumu$set$set_id == all_sets[i], ]
  p <- plot_cumu_set(dat_tmp, columns = 1, smooth = TRUE)
  plot_name = paste0("cumu_change_withLine_", all_sets[i], ".png")
  out_name <- paste(out_path, plot_name, sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, p, width = opts_long$width_short, height = opts_long$height_short, units = "in", scale = 0.8)
}
```


Graphs for each SET individually are not shown here but have been saved in _`r out_path`_


### Plus Sea Level Rise Line  

**This is an oversimplification of sea level rise**  

It is only meant to illustrate and give context for SET elevation change  



```{r, fig.width = opts_long$width_square, fig.height = opts_long$height_square}

# put slr slope into "per-day" rather than "per-year" because that's how R thinks
# 365.25 accounts for leap years
slr_for_graph <- slr/365.25

# generate a SET-specific intercept based on the first date of sampling at that SET
# also add in slope because geom_abline doesn't work otherwise
intercepts <- dat_cumu$set %>% 
  group_by(set_id) %>% 
  summarize(min_date = as.integer(min(date))) %>% 
  mutate(intercept = 0 - (slr_for_graph*min_date),
         slope = slr_for_graph)

# add the line to the plot
p <- plot_cumu_set(dat_cumu$set, columns = opts_long$multi_cols,
                   pointsize = 2,
                   smooth = FALSE)
p +
  geom_smooth(se = FALSE, method = 'lm', 
                    col = 'steelblue4') +
  geom_abline(data = intercepts,
              aes(intercept = intercept,
                  slope = slope),
              color = "red3",
              size = 1) +
  labs(subtitle = "red line has slope of long-term sea level rise; blue line is linear model of SET change")


# save it
out_name <- paste("R_output", "figures", "cumu_change_plots", "cumu_change_withLineAndSLR.png", sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, width = opts_long$width_square, height = opts_long$height_square, units = "in")
```

**The above graph is saved as:**   
`r out_name`   


```{r}
out_path <- paste("R_output", "figures", "cumu_change_plots", "individual_sets", sep = "/")

for(i in seq_along(all_sets)){
  dat_tmp <- dat_cumu$set[dat_cumu$set$set_id == all_sets[i], ]
  intercepts_tmp <- intercepts[intercepts$set_id == all_sets[i], ]
  
  
  p <- plot_cumu_set(dat_tmp, columns = opts_long$multi_cols,
                     pointsize = 2,
                     smooth = FALSE)
  p <- p +
    geom_smooth(se = FALSE, method = 'lm', 
                col = 'steelblue4') +
    geom_abline(data = intercepts_tmp,
                aes(intercept = intercept,
                    slope = slope),
                color = "red3",
                size = 1) +
    labs(subtitle = "red line has slope of long-term sea level rise; blue line is linear model of SET change")
  
  
  plot_name = paste0("cumu_change_withLineAndSLR", all_sets[i], ".png")
  out_name <- paste(out_path, plot_name, sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, p, width = opts_long$width_short, height = opts_long$height_short, units = "in", scale = 0.8)
}
```


Graphs for each SET individually are not shown here but have been saved in _`r out_path`_


```{r, fig.width = opts_long$width_square, fig.height = opts_long$height_square}
# add the line to the plot
p <- plot_cumu_set(dat_cumu$set, columns = opts_long$multi_cols,
                   smooth = FALSE,
                   pointsize = 2)
p + 
  geom_abline(data = intercepts,
              aes(intercept = intercept,
                  slope = slope),
              color = "red3",
              size = 0.8,
              alpha = 0.9) +
  labs(title = "Cumulative Change since first reading",
       subtitle = "red line has the slope of long-term sea level rise")


out_name <- paste("R_output", "figures", "cumu_change_plots", "cumu_change_withSLR.png", sep = "/")
out_file <- here::here(out_name)
ggsave(out_file, width = opts_long$width_square, height = opts_long$height_square, units = "in")
```


**The above graph is saved as:**   
`r out_name`   


```{r}
out_path <- paste("R_output", "figures", "cumu_change_plots", "individual_sets", sep = "/")

for(i in seq_along(all_sets)){
  dat_tmp <- dat_cumu$set[dat_cumu$set$set_id == all_sets[i], ]
  intercepts_tmp <- intercepts[intercepts$set_id == all_sets[i], ]
  
  
  p <- plot_cumu_set(dat_tmp, columns = opts_long$multi_cols,
                     pointsize = 2,
                     smooth = FALSE)

  p <- p + 
    geom_abline(data = intercepts_tmp,
                aes(intercept = intercept,
                    slope = slope),
                color = "red3",
                size = 0.8,
                alpha = 0.9) +
    labs(title = "Cumulative Change since first reading",
         subtitle = "red line has the slope of long-term sea level rise")

  plot_name = paste0("cumu_change_withSLR", all_sets[i], ".png")
  out_name <- paste(out_path, plot_name, sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, p, width = opts_long$width_short, height = opts_long$height_short, units = "in", scale = 0.8)
}
```


Graphs for each SET individually are not shown here but have been saved in _`r out_path`_



```{r, fig.width = opts_long$width_square, fig.height = opts_long$height_square}
# what if we just do it without any points; just smooths?
ggplot(dat_cumu$set, aes(x = date, y = mean_cumu)) +
        geom_line(col = 'gray70') +
        facet_wrap(~set_id, ncol = opts_long$multi_cols) +
        labs(title = 'Cumulative Change since first reading', 
             subtitle = 'light gray tracks mean change; blue is a linear smooth of change; \nred is slope of long-term sea level rise',
             x = 'Date',
             y = 'Change since first reading (mm)') +
        theme_classic() +
  geom_smooth(se = FALSE, method = 'lm', 
                    col = 'steelblue4',
              size = 1) +
  geom_abline(data = intercepts,
              aes(intercept = intercept,
                  slope = slope),
              color = "red3",
              size = 1)


out_name <- paste("R_output", "figures", "cumu_change_plots", "cumu_change_linesOnly.png", sep = "/")
out_file <- here::here(out_name)
ggsave(out_file, width = opts_long$width_square, height = opts_long$height_square, units = "in")
```


**The above graph is saved as:**   
`r out_name`   


```{r}
out_path <- paste("R_output", "figures", "cumu_change_plots", "individual_sets", sep = "/")

for(i in seq_along(all_sets)){
  dat_tmp <- dat_cumu$set[dat_cumu$set$set_id == all_sets[i], ]
  intercepts_tmp <- intercepts[intercepts$set_id == all_sets[i], ]

  p <- ggplot(dat_tmp, aes(x = date, y = mean_cumu)) +
    geom_line(col = 'gray70') +
    facet_wrap(~set_id, ncol = 1) +
    labs(title = 'Cumulative Change since first reading', 
         subtitle = 'light gray tracks mean change; blue is a linear smooth of change; \nred is slope of long-term sea level rise',
         x = 'Date',
         y = 'Change since first reading (mm)') +
    theme_classic() +
    geom_smooth(se = FALSE, method = 'lm', 
                col = 'steelblue4',
                size = 1) +
    geom_abline(data = intercepts_tmp,
                aes(intercept = intercept,
                    slope = slope),
                color = "red3",
                size = 1)

  
  plot_name = paste0("cumu_change_linesOnly", all_sets[i], ".png")
  out_name <- paste(out_path, plot_name, sep = "/")
  out_file <- here::here(out_name)
  
  ggsave(out_file, p, width = opts_long$width_square, height = opts_long$height_square, units = "in", scale = 0.8)

}
```


Graphs for each SET individually are not shown here but have been saved in _`r out_path`_



## Something similar, but with NAVD88  

Don't facet. Put all SETs in one facet. Use cumulative change, but instead of starting at "0" on the first date, add the NAVD88 elevation (either ground or receiver - all is relative so it doesn't really matter unless we want water level information, which involves another information "grab") to cumulative change data.  

```{r, fig.width = opts_long$width_tall, fig.height = opts_long$height_tall}
# if either column has NO NAs, use the column. Preference goes to ground elevation, then receiver elevation. 
# then move through partial scenarios: use ground elevation or receiver elevation, or make it NA.
navd88s <- mdat %>% 
  select(set_id = unique_set_id, 
         ends_with("navd88"))%>% 
  mutate_at(vars(matches("navd88")), as.numeric) %>% 
  mutate(set_navd88 = case_when(!anyNA(surface_elevation_ground_navd88) ~
                                 surface_elevation_ground_navd88,
                                !anyNA(surface_elevation_receiver_navd88) ~
                                 surface_elevation_receiver_navd88,
                                !is.na(surface_elevation_ground_navd88) ~
                                  surface_elevation_ground_navd88,
                                !is.na(surface_elevation_receiver_navd88) ~
                                    surface_elevation_receiver_navd88,
                                TRUE ~ 0))
         

cumu_navd88 <- left_join(dat_cumu$set, navd88s, by = "set_id") %>% 
  mutate(mean_cumu = (mean_cumu/1000) + set_navd88,
         set_id = fct_reorder(set_id, desc(set_navd88)))

ggplot(cumu_navd88, aes(x = date, y = mean_cumu, color = set_id)) +
  geom_line(size = 1) +
  theme_minimal() +
  labs(title = "SET cumulative change",
       subtitle = "first sample date set to that SET's NAVD88 elevation",
       y = "NAVD88 elevation (m)",
       color = "SET")


out_name <- paste("R_output", "figures", "cumu_change_plots", "cumu_change_NAVD88.png", sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, width = opts_long$width_tall, height = opts_long$height_tall, units = "in")
```


**The above graph is saved as:**   
_`r out_name`_   



# Only use sites with enough data  

From this point on, only SETs with 5 or more measurements over 4.5 or more years will be analyzed. If you do not have any SETs that have been measured for this long, you will NOT see analyses or graphs below.  

**Note** - if you're reading this at the SETr workshop, please note that I'll probably change the code some more and nothing below this section will show up if you don't have any sites with enough data to analyze.  

```{r}
# make a backup
dat_all <- dat

to_keep <- unlist(sample_info[sample_info$years_sampled >= 4.5 & sample_info$sample_events >= 5, "set_id"])

dat <- dat[dat$set_id %in% to_keep, ]

enough_dat <- nrow(dat) > 0
```


# Rate Calculations  

These rates were generated using linear mixed models. See Zuur et al. 2009 and Cahoon et al. 2019 for details.

Cahoon, D.R., Lynch, J.C., Roman, C.T. et al. Estuaries and Coasts (2019) 42: 1. https://doi.org/10.1007/s12237-018-0448-x  

Zuur, A.F., E.N. Ieno, N.J. Walker, A.A. Saveliev, and G.M. Smith. 2009. Mixed effects models and extensions in ecology with R. New York: Springer.  

The following is directly excerpted from Cahoon et al. 2019:

> Linear mixed models (LMMs, Zuur et al. 2009) were chosen
to analyze the surface elevation data. LMMs are ideal for
analyzing the nested longitudinal data that is produced by
the SET device. Rather than averaging the pin heights from
each SET prior to analysis, the measurements from each pin
are used as separate replicates. This preserves the variation
found within each SET and also maximizes statistical power.
Effects which are specific to each SET, direction within each
SET, and pin within each direction are treated as random
effects, which account for the lack of independence among
pins on the same SET. Analysis was performed using
mixed-effect models in the nlme package (Pinheiro et al.
2016) in R version 3.3.2 (R Core Team 2016).
Data from each of the five sites were analyzed separately.
Pin height (m, NAVD88) served as the response variable, and
the fixed effects were the number of days since the initial
reading was taken. To account for a potential reduction of
independence among pins on the same SET, the model included
a random slope and intercept for each pin, nested in the SET
position on the benchmark (typically, four positions were read
during each sampling event), nested in the SET. This model
was first fit using maximum likelihood. It was then compared
to a model with identical random effects but an intercept-only
fixed effect using the corrected form of Akaike’s information
criterion (AICc, Akaike 1974, Burnham and Anderson 2004).
If the intercept-only model was superior, this indicated that
there is no trend in elevation over time. The model including
a trend through time was then refit using restricted maximum
likelihood to estimate the coefficients of the regression. For
comparative purposes, this was done even in cases where the
intercept-only model was superior.


In our case, we have, for each SET:  

+  **response variable:** pin_height  
+  **fixed effect:** date  
+  **random effects:** arm_position, pin_number  (note, these are nested)  

## Rates of change  

These rates were generated using the `lme()` function in the `nlme` package. Confidence intervals were generated using the `intervals()` function, also in the `nlme` package. 

All calculations generated output in *mm/day* and were converted to *mm/yr* by multiplying by 365.25 (this accounts for leap years).    

```{r calc-lmm, eval = enough_dat}
models2 <- dat %>%
    group_by(reserve, set_id) %>%
    do(mod = lme(pin_height ~ date, data = ., random = ~1|arm_position/pin_number, na.action = na.omit))

lmm_out <- models2 %>% 
  mutate(rate = intervals(mod, which = "fixed")$fixed["date", 2] * 365.25,
         CI_low = intervals(mod, which = "fixed")$fixed["date", 1] * 365.25,
         CI_high = intervals(mod, which = "fixed")$fixed["date", 3] * 365.25) %>% 
  select(-mod)

lmm_out %>% 
  flextable() %>% 
  theme_booktabs() %>% 
  autofit()
```



### Additional model diagnostics  

```{r addl-diagnostics}
diagnostics_out <- models2 %>% 
  mutate(sigma = sigma(mod),
         AIC = AIC(mod),
         BIC = BIC(mod),
         logLik = mod$logLik) %>% 
  select(-mod)

diagnostics_out %>% 
  flextable() %>% 
  theme_booktabs() %>% 
  autofit()
```

### Combine information from rate calculations, metadata, and SLR into one file  

```{r combine-rates-mdat-slr, eval = enough_dat}
# pick out relevant fields from the metadata to be used for mapping
mdat_sub <- mdat %>% 
  select(unique_set_id, user_friendly_set_name, reserve, latitude_dec_deg, longitude_dec_deg, co_dominant_species1)

# join that to the rates generated by the model output above
rate_summary <- left_join(lmm_out, mdat_sub, by = c("reserve" = "reserve", "set_id" = "unique_set_id"))

# now add in slr info, and also calculate the lower and upper CI limits for SLR
rate_summary <- left_join(rate_summary, slr_rates, by = c("reserve" = "reserve")) %>% 
  rename(slr_rate = slr_rate_mm_yr) %>% 
  mutate(slr_CI_low = slr_rate - x95_percent_ci,
         slr_CI_high = slr_rate + x95_percent_ci,
         set_slr_ratio = rate / slr_rate) %>% 
  select(reserve, set_id, rate, CI_low, CI_high, slr_rate, slr_CI_low, slr_CI_high, everything())

# write that to a csv for use in mapping script
write.csv(rate_summary, 
          here::here("data", "intermediate", "rate_summary.csv"),
          row.names = FALSE)
```




```{r slr-comp-assignment, eval = enough_dat}
rates_slr_comp <- rate_summary %>%
    mutate(dir_0 = case_when(CI_high < 0 ~ "dec_sig",
                             CI_low > 0  ~ "inc_sig",
                             rate < 0 ~ "dec_nonsig",
                             rate > 0 ~ "inc_nonsig",
                             TRUE ~ "nonsig"),
           dir_slr = case_when(CI_high < slr_CI_low ~ "dec_sig",
                               CI_low > slr_CI_high  ~ "inc_sig",
                               rate < slr_rate ~ "dec_nonsig",
                               rate > slr_rate ~ "inc_nonsig",
                               TRUE ~ "nonsig")) %>% 
  select(reserve, set_id, rate, CI_low, CI_high, set_slr_ratio, dir_0, dir_slr)
```

***
***


# Increasing/Decreasing (Comparison to 0)  

The following tables break the SETs into groups where the rate of SET elevation change is *lower than* / *higher than* / *not different from* 0. *Lower than* and *higher than* tables imply that the 95% confidence intervals for the SET's rate of change do not include 0. *Not different from* means that 0 *is* included.  


<br>

***

## SET Elevation Change < 0 mm/yr  


```{r zero-comp-lower, eval = enough_dat}
rates_slr_comp %>%
    filter(dir_0 == "dec_sig") %>%
    select(-dir_0, -dir_slr, -set_slr_ratio) %>% 
    flextable() %>% 
    theme_booktabs() %>% 
    autofit()
```

<br>

***

## SET Elevation Change > 0 mm/yr  


```{r zero-comp-higher, eval = enough_dat}
rates_slr_comp %>%
    filter(dir_0 == "inc_sig") %>%
    select(-dir_0, -dir_slr, -set_slr_ratio) %>% 
    flextable() %>% 
    theme_booktabs() %>% 
    autofit()
```

<br>

***

## SET Elevation Change 95% CI Includes 0 mm/yr  


```{r zero-comp-same, eval = enough_dat}
rates_slr_comp %>%
    filter(dir_0 %in% c("dec_nonsig", "inc_nonsig")) %>%
    select(-dir_0, -dir_slr, -set_slr_ratio) %>% 
    flextable() %>% 
    theme_booktabs() %>% 
    autofit()
```


***
***


# Sea Level Rise Comparisons  

The long-term local rate of sea level rise is **`r slr` +/- `r slr_ci` mm/yr** .  

This rate is reported by `r slr_res$nearest_nwlon_station`, NWLON station number `r as.character(slr_res$nwlon_station_number)` based on data from *`r slr_res$data_start`* to *`r slr_res$data_end`*.  

The following tables break the SETs into groups where the rate of SET elevation change is *lower than* / *higher than* / *not different from* this SLR rate. *Lower than* and *higher than* tables imply that 95% confidence intervals do not overlap between the SET and SLR. *Not different from* means that confidence intervals *do* overlap.  

 
<br>

<br>

***

## SET Elevation Change < SLR; CIs don't overlap  


```{r slr-comp-lower, eval = enough_dat}
rates_slr_comp %>%
    filter(dir_slr == "dec_sig") %>%
    select(-dir_0, -dir_slr) %>% 
    flextable() %>% 
    theme_booktabs() %>% 
    autofit()
```

<br>


***

## SET Elevation Change > SLR; CIs don't overlap


```{r slr-comp-higher, eval = enough_dat}
rates_slr_comp %>%
    filter(dir_slr == "inc_sig") %>%
    select(-dir_0, -dir_slr) %>% 
    flextable() %>% 
    theme_booktabs() %>% 
    autofit()
```

<br>

***

## SET Elevation Change and SLR CIs overlap  


```{r slr-comp-same, eval = enough_dat}
rates_slr_comp %>%
    filter(dir_slr %in% c("dec_nonsig", "inc_nonsig")) %>%
    select(-dir_0, -dir_slr) %>% 
    flextable() %>% 
    theme_booktabs() %>% 
    autofit()
```


***
***

Add dominant veg to model_out data frame for graphing purposes

```{r add-veg}
lmm_out <- left_join(lmm_out, select(mdat, unique_set_id, co_dominant_species1), by = c("set_id" = "unique_set_id")) %>% 
    rename(veg = co_dominant_species1)
```


## Graphical Comparison to Sea Level Rise and 0  


```{r plot-slr-ci-lmm, fig.width = opts_long$width_short, fig.height = opts_long$height_short, eval = enough_dat}

plot_rate_comps(data = lmm_out, plot_type = 3, set_ids = set_id,
                set_ci_low = CI_low, set_ci_high = CI_high,
                rates = rate, slr = slr, slr_ci = slr_ci)


out_name <- paste("R_output", "figures", "summary_plots", "summary_plot.png", sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, width = opts_long$width_short, height = opts_long$height_short, scale = 0.8)

cat(paste("The above graph is saved as: ", out_name))
```



```{r plot-slr-ci-lmm-veg, fig.width = opts_long$width_short, fig.height = opts_long$height_short, eval = enough_dat}
# colored by veg
plot_rate_comps(data = lmm_out, plot_type = 3, set_ids = set_id,
                set_ci_low = CI_low, set_ci_high = CI_high,
                rates = rate, slr = slr, slr_ci = slr_ci,
                color_by_veg = TRUE, veg = veg)

out_name <- paste("R_output", "figures", "summary_plots", "summary_plot_veg.png", sep = "/")
out_file <- here::here(out_name)

ggsave(out_file, width = 1.2*opts_long$width_short, height = opts_long$height_short, scale = 0.8)

cat(paste("The above graph is saved as: ", out_name))
```



***

***

In order, according to specifications in the metadata; and labeled with user-friendly names. Note that no plot will be produced if there are any NAs in the metadata fields `numerical_order` or `user_friendly_set_name`.    

```{r plot-slr-ci-lmm-ordered, fig.width = opts_long$width_short, fig.height = opts_long$height_short, eval = enough_dat}
# only run this chunk if numerical order and user friendly names are specified for all SETs

# friendly names need more space - up width in ggsave by 10%
# veg legends need more space - up width in ggsave by 20%
# if both: 25% wider???

if(sum(is.na(mdat$numerical_order)) + sum(is.na(mdat$user_friendly_set_name)) == 0){
  # make the set id a factor
  lmm_out_ordered <- mdat %>% 
    select(unique_set_id, user_friendly_set_name, numerical_order) %>%
    left_join(lmm_out, ., by = c("set_id" = "unique_set_id")) 
  
  a <- forcats::fct_reorder(lmm_out_ordered$user_friendly_set_name,
                            desc(lmm_out_ordered$numerical_order))
  b <- forcats::fct_reorder(lmm_out_ordered$set_id,
                            desc(lmm_out_ordered$numerical_order))
  
  lmm_out_ordered$user_friendly_set_name <- a
  lmm_out_ordered$set_id2 <- b
  
  
  p <- plot_rate_comps(data = lmm_out_ordered, plot_type = 3, 
                       set_ids = set_id2, set_ci_low = CI_low,
                       set_ci_high = CI_high, slr = slr, slr_ci = slr_ci,
                       rates = rate) 
  
  print(p)
  
  out_name <- paste("R_output", "figures", "summary_plots", "summary_plot_ordered.png", sep = "/")
  out_file <- here::here(out_name)
  
  ggsave(out_file, p, width = opts_long$width_short, height = opts_long$height_short, scale = 0.8)
  
  cat(paste("The above graph is saved as: ", out_name))
  
  
  
  p <- plot_rate_comps(data = lmm_out_ordered, plot_type = 3, 
                       set_ids = set_id2, set_ci_low = CI_low,
                       set_ci_high = CI_high, slr = slr, slr_ci = slr_ci,
                       rates = rate, color_by_veg = TRUE, veg = veg) 
  
  print(p)
  
  out_name <- paste("R_output", "figures", "summary_plots", "summary_plot_ordered_veg.png", sep = "/")
  out_file <- here::here(out_name)
  
  ggsave(out_file, p, width = 1.1*opts_long$width_short, height = opts_long$height_short, scale = 0.8)
  
  cat(paste("The above graph is saved as: ", out_name))
  
  
  
  p <- plot_rate_comps(data = lmm_out_ordered, plot_type = 3, 
                       set_ids = user_friendly_set_name, set_ci_low = CI_low,
                       set_ci_high = CI_high, slr = slr, slr_ci = slr_ci,
                       rates = rate) 
  
  print(p)
  
  out_name <- paste("R_output", "figures", "summary_plots", "summary_plot_ordered_friendly_names.png", sep = "/")
  out_file <- here::here(out_name)
  
  ggsave(out_file, p, width = opts_long$width_short, height = opts_long$height_short, scale = 0.8)
  
  cat(paste("The above graph is saved as: ", out_name))
  
  
  
  p <- plot_rate_comps(data = lmm_out_ordered, plot_type = 3, 
                       set_ids = user_friendly_set_name, set_ci_low = CI_low,
                       set_ci_high = CI_high, slr = slr, slr_ci = slr_ci,
                       rates = rate, color_by_veg = TRUE, veg = veg) 
  
  print(p)
  
  out_name <- paste("R_output", "figures", "summary_plots", "summary_plot_ordered_friendly_names_veg.png", sep = "/")
  out_file <- here::here(out_name)
  # because the legend is pretty wide for the plots with veg coloring, I'm saving them 20% wider than specified. this makes the final dimensions of the main panel comparable to the non-veg plots.
  ggsave(out_file, p, width = 1.1*opts_long$width_short, height = opts_long$height_short, scale = 0.8)
  
  cat(paste("The above graph is saved as: ", out_name))
}
```


***

## MAPS  

```{r map-script-source, eval = enough_dat}
source(here::here("R_scripts", "sourced", "006_map_making_static.R"))
```


### Comparisons to 0  

```{r map0, eval = enough_dat}
m0
```

### Comparisons to SLR  

```{r mapSLR, eval = enough_dat}
mSLR
```

